Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib

### 2023-08-21 supplement/figure-env/figure-example.tex

\abstract{
This is a small sample article to demonstrate usage of texor to convert figure environments.
}

@@

Images are an essential component in any article, However due to the differences in the support for various graphic formats between LaTeX and markdown/HTML we need to fallback on raster graphics. 
It is well summarized in the Table \ref{table:1}.

@@

\section{Image with width parameters}
This section may contain a figure such as Figure~\ref{figure:rlogo}.
This is the most basic example of figure.

@@

@@

Image \ref{fig:normal} is a graphical representation of normal distribution.

@@

\section{Multiple images}
 Pandoc v3 and above now supports a new Figure object \citep{pandoc} which supports multiple 
 images side by side or in a grid format.
%% Two images side by side
\subsubsection{Two or more Images side by side}
\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure}
\end{verbatim}

@@

Here is a tikz image example in fig \ref{fig:tikz} adapted from \citep{casflow}.

Another interesting aspect of including tikz image here is that you can modify the
source and re-convert without making any other changes. Figure will get updated
in the generated html article.

\subsubsection{Tikz Code:}
The figure  \ref{fig:tikz} is an example of how \pkg{texor} handles tikz images,

\begin{itemize}
\item First figures containing tikz images are isolated.
\item Tikz libraries are fetched from the wrapper file,
\item Tikz Code section is isolated into a standalone LaTeX file and compiled.
\item The Compiled LaTeX file generates a PDF Image.
\item This is then Converted again to PNG format as PDFs embedded in HTML wont look good.
\item In the figure environment a line is added \verb|\includegraphics{tikz/somefigure.png}|
\item During the conversion in pandoc, a lua filter removes extra/unncessary text from the rest of the figure
environemnt.
\end{itemize}

If You use \pkg{texor} to convert your articles using \verb|texor::latex_to_web()| with \verb|temp_mode=TRUE|(it is TRUE by default). The resultant Rmarkdown/HTML file will not modify the contents of your LaTeX file. In this case you can keep reloading the article after making changes to the tikz images, without having to do the above steps manually in case you are converting the article by hand.

@@

we do support algorithm2e diagrams and images, these will be numbered differently
and we strongly suggest to use "alg:" in labels for best results. 
Here is an example of algorithm \ref{alg:how}, referenced from \cite{algoexample}

@@

Figures can also house non-image environments like CodeBlocks and WideTables. This is 
the part where things start to look a bit different in the converted articles, CodeBlocks 
in figure environments would have numbering and caption that goes "Figure x:". However, 
when the same CodeBlock is converted to Rmarkdown, they are numbered as "CodeBlock x:".
This distinguishes them from normal Images and matches with authors vision more appropriately.

As for Inclusion of widetables, since it houses two different tabular environments with the same caption,
it would be treated as two different tables by pandoc sharing the same caption. This would not be the best solution hence, \pkg{texor} package uses Figure enviornment to house the widetables.
This way they are distinguished well from normal tables, also the numbering scheme is seperated from figures "WideTable x:".

@@

\begin{figure}[htbp]
\begin{center}
\begin{verbatim}
code_in_figure <- function() {
  if (pandoc_version >= 3) {
    print("Code in Figure Supported")
  }
  else {
    print("code in Figure not supported")
  }
}
\end{verbatim}
\caption{ Example Code inside Figure environment}
\label{code:example}
\end{center}
\end{figure}
\section{Summary}

>>>>

\abstract{
This is a small sample article to demonstrate usage of \CRANpkg{texor} to convert figure environments.
}

@@

Images are an essential component of any article, however, due to the differences in support for various graphic formats between LaTeX and markdown/HTML we need to fall back on raster graphics. 
The support for different image formats across markup languages is summarized in Table~\ref{table:1}.

@@

\section{Image with width parameters}

The following code includes an image with width parameters, producing the output in Figure~\ref{figure:rlogo}.

@@

This is the most basic example of figure.

@@

The following code includes an image in PDF format, shown in Figure \ref{fig:normal}.

@@

\section{Multiple images}
 Pandoc v3 and above now support a new \code{Figure} object \citep{pandoc} which supports multiple 
 images side by side or in a grid format. The following subsections demonstrate these capabilities.
%% Two images side by side
\subsubsection{Two or more Images side by side}
\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure}
\end{verbatim}

@@

Figure \ref{fig:tikz} shows a tikz image adapted from \citep{casflow}.

An interesting aspect of including a tikz image here is that you can modify the
source and re-convert without making any other changes. Figure will get updated
in the generated html article.

\subsubsection{Tikz Code:}
The image in Figure  \ref{fig:tikz} is a graphical representation of how \pkg{texor} handles tikz images:

\begin{itemize}
\item Figures containing tikz images are isolated.
\item Tikz libraries are fetched from the wrapper file.
\item The Tikz code section is isolated into a standalone LaTeX file and compiled.
\item The compiled LaTeX file generates a PDF Image.
\item This is converted to PNG format for embedding in the HTML output.
\item \verb|\includegraphics{tikz/somefigure.png}| is added to the figure environment.
\item A lua filter removes redundant text the figure environment.
\end{itemize}

If you use \pkg{texor} to convert your articles using \verb|texor::latex_to_web()| with \verb|temp_mode=TRUE|(it is \verb|TRUE| by default). The resultant Rmarkdown/HTML file will not modify the contents of your LaTeX file. In this case you can keep reloading the article after making changes to the tikz images, without having to do the above steps manually in case you are converting the article by hand.

@@

Diagrams and images using the \code{algorithm2e} environment are supported, these will be numbered differently.
We strongly suggest to use "alg:" in labels for best results. 
Algorithm \ref{alg:how} is an example from the \CRANpkg{algorithm2e} vignette \cite{algoexample}.

@@

Figures can also house non-image environments like code blocks and wide tables. This is 
one aspect where the converted articles differ between PDF and HTML formats. Code blocks 
in figure environments would have numbering and caption of the form "Figure x:" in the PDF output. However, 
when the code blocks are converted to R Markdown, they are numbered as "CodeBlock x:".
This distinguishes them from normal images and matches with the authors' vision more appropriately.

As for the inclusion of wide tables, since the figure houses two different tabular environments with the same caption,
this would be treated by pandoc as two different tables sharing the same caption. This would not be the best solution hence, the \pkg{texor} package uses Figure environment to house the widetables.
This way they are distinguished well from normal tables, also the numbering scheme is seperated from figures "WideTable x:".

@@

\begin{figure}[htbp]
\begin{center}
\begin{verbatim}
code_in_figure <- function() {
  if (pandoc_version >= 3) {
    print("code in figure supported")
  }
  else {
    print("code in figure not supported")
  }
}
\end{verbatim}
\caption{ Example Code inside Figure environment}
\label{code:example}
\end{center}
\end{figure}
\section{Summary}

### 2023-08-21 supplement/math-env/math-example.tex

\abstract{
This is a small sample article to demonstrate usage of texor to convert math environments.
}

\section{Introduction}
Math typesetting has always been LaTeX’s highlight feature, making it a de facto choice among academicians and researchers globally. However, as we proceed to our humble web interfaces, math is hard to describe traditionally. There have been advancements in JavaScript libraries to better Typeset and present math in web pages but not all LaTeX commands/math functions are available.

@@

\subsubsection{Pandoc Handling, extensions}
As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the Rmarkdwon file into HTML,
the \verb|rjtools::rjournal_web_article| template by default uses MathJax as the math engine. 
Also we specify the Mathjax version in the metadata of the generated Rmarkdown file.

\section{Inline math}
One can define Inline Math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
It is also possible to use inline math within captions and  
\begin{table}[htbp]
  \centering
  \begin{tabular}{| c | c |}
  \hline
  Command & Math\\
    \hline
    \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
\(\mu = (0,0,0)^\top \)
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    \( \mu = (0,0,0)^\top \)
    \end{minipage}\\
    \hline

  \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
$\mu = (0,0,0)^\top $
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    $ \mu = (0,0,0)^\top $
    \end{minipage}\\
    \hline
  \end{tabular}
  \caption{InlineMath syntax and its output side by side}
  \label{table:1}
\end{table}

\section{Display Math}

Display Math refers to equations, almost all LaTeX equations can be rendered by
MathJax.

\begin{verbatim}
\begin{align}
f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
  \label{eq:1}
\end{align}
\end{verbatim}
This will render as \ref{eq:1}.

@@

\section{Equation Numbering}
In LaTeX your equations get numbered automatically unless you are using \verb|\[..\]| environment or \verb|\nonumber| to suppress
numbering. Equation numbering works a bit differently in \CRANpkg{bookdown} (the base of RJ web article format) where
it is mandatory to have a \verb|(\#eq:xx)| which is described in more detail in \citep{bookdown}.

The \CRANpkg{texor} package eases authors from manually adding \verb|(\#eq:xx)| to equations in Rmarkdown
by using pandoc Lua filter to convert existing \verb|\label{..}| in the equations to \verb|(\#eq:xx)|
during conversion. Equation \ref{eq:xe} shows such a use case of equation being numbered by its label in Rmakrdown
as well as LaTeX.

@@

\section{Custom Math Commands}

%Also avoid custom math commands MathJaX will never be able to parse it. The reason being
%MathJax renders math case by case and custom commands will already be removed by pandoc
%during the conversion, hence MathJax will not be aware of such redefinations.
To ease the frequent use of the command, some custom math commands that contain LaTeX math commands. 
However, This will not function if you use functions/environments from CTAN packages or 
the ones that are out-of-the-box functions which do not render in HTML/markdown.

\paragraph{Acceptable}

@@

\paragraph{Not acceptable}

@@

\section{Unsupported LaTeX Commands}

Altough MathJax does a good job of supporting most LaTeX math functions,
some functions do not work like \verb|\bm| \verb|\boldmath|, instead use \verb|\mathbf|.

@@

In summary the \CRANpkg{texor} package with the help of pandoc and MathJax supports:
\begin{itemize}
\item some common math environments.
\item Equation Numbering
\end{itemize}

>>>>

\abstract{
This is a small sample article to demonstrate usage of the texor package to convert math environments.
}

\section{Introduction}
Math typesetting has always been a highlight of LaTeX, making it a de facto choice among academics and researchers globally. However, in the move to humble web interfaces, math has been traditionally hard to describe. There have been advances in JavaScript libraries to better typeset and present math in web pages but not all LaTeX commands/math functions are available.

@@

\subsubsection{Pandoc handling, extensions}
As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the R Markdown file into HTML,
the \verb|rjtools::rjournal_web_article| template uses MathJax as the math engine by default. 
We also specify the Mathjax version in the metadata of the generated Rmarkdown file.

\section{Inline math}
One can define inline math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
It is also possible to use inline math within captions and  
\begin{table}[htbp]
  \centering
  \begin{tabular}{| c | c |}
  \hline
  Command & Math\\
    \hline
    \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
\(\mu = (0,0,0)^\top \)
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    \( \mu = (0,0,0)^\top \)
    \end{minipage}\\
    \hline

  \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
$\mu = (0,0,0)^\top $
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    $ \mu = (0,0,0)^\top $
    \end{minipage}\\
    \hline
  \end{tabular}
  \caption{Inline math syntax and its output side by side.}
  \label{table:1}
\end{table}

\section{Display math}

Display math refers to equations typeset on separate lines rather than in line with the text. Almost all LaTeX equations can be rendered by
MathJax.

\begin{verbatim}
\begin{align}
f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
  \label{eq:1}
\end{align}
\end{verbatim}
The code above will render as Figure~\ref{eq:1}.

@@

\section{Equation numbering}
In LaTeX your equations get numbered automatically unless you are using \verb|\[..\]| environment or \verb|\nonumber| to suppress
numbering. Equation numbering works a bit differently in \CRANpkg{bookdown} (the base of the R Journal web article format) where
it is mandatory to have a \verb|(\#eq:xx)| which is described in more detail in \citet{bookdown}.

The \CRANpkg{texor} package relieves authors from manually adding \verb|(\#eq:xx)| to equations in R Markdown
by using a pandoc Lua filter to convert existing \verb|\label{..}| in the equations to \verb|(\#eq:xx)|
during conversion. Equation~\ref{eq:xe} shows such a use case of an equation being numbered by its label in R Markdown
as well as LaTeX.

@@

\section{Custom math commands}

%Also avoid custom math commands MathJaX will never be able to parse it. The reason being
%MathJax renders math case by case and custom commands will already be removed by pandoc
%during the conversion, hence MathJax will not be aware of such redefinations.
To ease the frequent use of the command, some custom math commands that contain LaTeX math commands. 
However, this will not work if you use functions/environments from CTAN packages or 
the ones that are out-of-the-box functions which do not render in HTML/markdown.

\paragraph{Renders correctly in HTML}

@@

\paragraph{Does not render correctly in HTML}

@@

\section{Unsupported LaTeX commands}

Altough MathJax does a good job of supporting most LaTeX math functions,
some functions do not currently work. Common examples are \verb|\bm| and \verb|\boldmath|, for which \verb|\mathbf| can be used instead.

@@

In summary the \CRANpkg{texor} package, with the help of pandoc and MathJax supports:
\begin{itemize}
\item Common math environments.
\item Equation numbering.
\end{itemize}

### 2023-08-21 supplement/table-env/table-example.tex
almost first draft

\section{Introduction}
Tables are commonly used in RJ articles to display data in a tabular format. However, there are differences in the way tables are handled by LaTeX and HTML.
LaTeX tables have more customization and are usually optimized for printing, whereas the web articles need tables optimized for varying sizes of media.
Pandoc converts most of the tables somewhat easily, but is unable to do well with table customization packages and complex tables.
Some pandoc extensions are used in order to tackle them, they are : \verb|simple_tables|,\verb|pipe_tables|
\section{Generic tables}
Normal LaTeX tables will be converted just fine, even if they include a bit of math and other elements.
However do note that the conversion will be done to traditional or pandoc style markdown instead of Rmarkdown
specific using \verb|knitr::kable(..)| function. Also any custom LaTeX commands or specific font charactars would
not be supported.
\begin{table}[htbp]
\centering
\begin{tabular}{l | llll }
 \hline
 Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
 \hline
 PNG       & Yes & Yes & Yes & Yes \\
 JPG       & Yes & Yes & Yes & Yes \\
 PDF       & Yes & No & No & No \\
 SVG       & No & Yes & Yes & Yes \\
 Tikz      & Yes & No & Yes & No \\
 Algorithm & Yes & No & No & No \\
\hline
\end{tabular}
\caption{Image Format support in various Markup/Typesetting Languages}
\label{table:1}
\end{table}

\section{Multicolumn tables}
\begin{table*}[htbp]
\begin{center}
\begin{tabular}{l | llll }
 \hline
 \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
\multicolumn{2}{c}{$Y$} \\
 \hline
  & 1 & 2 & 1 & 2 \\
 EX1  & X11 & X12 &  Y11  & Y12 \\
 EX2  & X21 & X22 &  Y21  & Y22 \\
 EX3  & X31 & X32 &  Y31  & Y32 \\
 EX4  & X41 & X42 &  Y41  & Y42\\
 EX5  & X51 & X52 &  Y51  & Y52 \\
\hline
\end{tabular}
\caption{An Example Table with Multicolumn}
\label{table:2}
\end{center}
\end{table*}

The above example illustrates a multicolumn table \ref{table:2}, which the \CRANpkg{texor} package can handle through subroutines and pre-processing step to transform the LaTeX code.

@@

Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the markdown/HTML will not support the specific changes table* environment will bring.

\section{Complex tables}
A complex table with various other elements like figures, math , code and so on, are also supported
by \CRANpkg{texor}.  
\begin{table}[htbp]
\centering
\begin{tabular}{|l | lll |}
 \hline
 Inline Format & LaTeX Support & Web Support &  Rendering \\
 \hline
 Text       & Yes & Yes  & \emph{Hello} \\ \hline
 Image       & Yes & Yes & \begin{minipage}{0.25\textwidth}
    \centering
    \includegraphics[width=1\textwidth]{Rlogo-5.png}
    \end{minipage} \\ \hline
 CodeBlock      & Yes & Yes & \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
x <- 1:100
y <- dbinom(x,100,prob = 0.5)
plot(x,y)
\end{example}
    \end{minipage} \\ \hline
 Math      & Yes & Yes & $e = m c^2 $ \\ \hline
 Link      & Yes & Yes &  \href{www.google.com}{Google} \\ \hline
 Nested Table & Yes & No & NaN \\
\hline
\end{tabular}
\caption{Image Format support in various Markup/Typesetting Languages}
\label{table:3}
\end{table}

@@

Widetable environments also gets some perks when converted by \CRANpkg{texor} package. 
They get their own block and seperate numbering (although references are shared by figures).

@@

Here is a reference to the table \ref{table:4} and table \ref{table:5}.
\section{Summary}

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item some common table environments.
\item wide tables.
\item some multicolumn tables.
\item inclusion of different environments like figure/code in tables.
\end{itemize}

>>>>

\section{Introduction}
Tables are commonly used in R Journal articles to display data in a tabular format. However, there are differences in the way tables are handled by LaTeX and HTML.
LaTeX tables have more customization and are usually optimized for printing, whereas the web articles need tables optimized for varying sizes of media.
Pandoc converts most of the tables somewhat easily, but is unable to do well with table customization packages and complex tables.
The \CRANpkg{texor} packages uses the pandoc extensions \verb|simple_tables| and \verb|pipe_tables| to tackle these difficult cases.

\section{Generic tables}
Normal LaTeX tables will be converted just fine, even if they include a some math and other elements.
However, note that the conversion will be done to traditional or pandoc-style markdown rather than R Markdown with the \verb|knitr::kable(..)| function. Also any custom LaTeX commands or specific font characters will
not be supported.
\begin{table}[htbp]
\centering
\begin{tabular}{l | llll }
 \hline
 Graphics Format & LaTeX & Markdown & R Markdown & HTML \\
 \hline
 PNG       & Yes & Yes & Yes & Yes \\
 JPG       & Yes & Yes & Yes & Yes \\
 PDF       & Yes & No & No & No \\
 SVG       & No & Yes & Yes & Yes \\
 Tikz      & Yes & No & Yes & No \\
 Algorithm & Yes & No & No & No \\
\hline
\end{tabular}
\caption{Image format support in various markup/typesetting languages.}
\label{table:1}
\end{table}

\section{Multicolumn tables}
\begin{table*}[htbp]
\begin{center}
\begin{tabular}{l | llll }
 \hline
 \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
\multicolumn{2}{c}{$Y$} \\
 \hline
  & 1 & 2 & 1 & 2 \\
 EX1  & X11 & X12 &  Y11  & Y12 \\
 EX2  & X21 & X22 &  Y21  & Y22 \\
 EX3  & X31 & X32 &  Y31  & Y32 \\
 EX4  & X41 & X42 &  Y41  & Y42\\
 EX5  & X51 & X52 &  Y51  & Y52 \\
\hline
\end{tabular}
\caption{An example table using \code{multicolumn}}
\label{table:2}
\end{center}
\end{table*}

Table~\ref{table:2} illustrates a table that uses the \code{multicolumn} command, which the \CRANpkg{texor} package can handle through subroutines and pre-processing steps to transform the LaTeX code.

@@

Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.

\section{Complex tables}
A complex table with various other elements like figures, math, code and so on, are also supported
by \CRANpkg{texor}.  
\begin{table}[htbp]
\centering
\begin{tabular}{|l | lll |}
 \hline
 Inline Format & LaTeX Support & Web Support &  Rendering \\
 \hline
 Text       & Yes & Yes  & \emph{Hello} \\ \hline
 Image       & Yes & Yes & \begin{minipage}{0.25\textwidth}
    \centering
    \includegraphics[width=1\textwidth]{Rlogo-5.png}
    \end{minipage} \\ \hline
 CodeBlock      & Yes & Yes & \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
x <- 1:100
y <- dbinom(x,100,prob = 0.5)
plot(x,y)
\end{example}
    \end{minipage} \\ \hline
 Math      & Yes & Yes & $e = m c^2 $ \\ \hline
 Link      & Yes & Yes &  \href{www.google.com}{Google} \\ \hline
 Nested Table & Yes & No & NaN \\
\hline
\end{tabular}
\caption{Image format support in various markup/typesetting languages}
\label{table:3}
\end{table}

@@

Widetable environments are treated specially when converted by the \CRANpkg{texor} package. 
They get their own block and separate numbering (although references are shared by figures).

@@

Here is a reference to Table~\ref{table:4} and Table~\ref{table:5}.
\section{Summary}

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item Some common table environments.
\item Wide tables.
\item Some tables with \code{multicolumn} commands.
\item Environments such as \code{figure}/\code{code} in tables.
\end{itemize}

### 2023-08-21 texor.Rmd

The R Journal is the primary open-access outlet for publications produced by the R community. It was started in 2009, evolving from a newsletter that ran from 2001, to become a more formal publication that encourages and provides credit for the documentation of statistical computing research.

>>>>

The R Journal is the primary open-access outlet for publications produced by the R community. It was started in 2009, evolving from the R News newsletter that ran from 2001, to become a more formal publication that encourages and provides credit for the documentation of statistical computing research.
